\pdftitle

### 2017-05-17 maxNt.tex
new material on projection

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-17T00:18:49+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

We have a population of $N$ binary neurons. We assume that they can be
distinguished, by their spike shapes for example; but other details, like
their locations, are unknown. The neurons have a joint state
$(X_1,\dotsc,X_N) \defs \yX$ having fixed but unknown binary values
$(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$. A particular sample of $n$
neurons from this population has joint state $(x_1, \dotsc, x_n) \defs \yx$
having fixed binary values $(r_1, \dotsc, r_n)\defs\yr$. We will consider
various averages of the population and the sample. For this purpose we
introduce a general averaging operator $\sav{\dotv}$ defined by
\begin{align}
  \av{\yX}  &\defd \tfrac{1}{N} (X_1 + X_2 + \dotsb + X_N),
  &
  \sav{\yx}  &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
    \\[2\jot]
  \av{\yX \yX} &\defd
\tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3  + \dotsb +  X_{N-1} X_N),
  &
    \sav{\yx \yx} &\defd
\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3  + \dotsb +  x_{n-1} x_n),
\\[2\jot]
\av{\yX\yX\yX} &\defd
                  \tbinom{N}{3}^{-1}
                  (X_1 X_2 X_3 +% X_1 X_2 X_4
                   \dotsb + X_{N-2} X_{N-1}X_N),
  &
    \sav{\yx\yx\yx} &\defd
                      \tbinom{n}{3}^{-1} (x_1 x_2 x_3 + %x_1 x_2 x_4
                       \dotsb + x_{n-2} x_{n-1} x_n),
\end{align}
and so on; analogously for the quantities like $\yRf$ and $\sav{\yr\yr}$.
These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
the fraction of simultaneously active pairs out of all $\binom{N}{2}$
pairs, $\av{\yX\yX\yX}$ the fraction of simultaneously active triplets, and
so on. Products of states like $X_i \dotsm X_j$ also have values in
$\set{0,1}$; from this it can combinatorially be proved that
\begin{equation}
  \label{eq:products_intermsof_average}
  \av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}}
  = \binom{N}{m}^{-1}\binom{N\yXf}{m}
\end{equation}
and similarly for $\yx$, $\yR$, $\yr$.

@@

By marginalization we obtain the probability for the state of the sample:
\begin{gather}
  \label{eq:marginal}
  \p(\yx = \yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs \| \yH),
  \\
  \shortintertext{with}
  \label{eq:subpop_average}
  \p(\yxs=\yrs \| \yH) = \sum_{N \yRf =0}^{N}
    \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
  \p(\yXf=\yRf \| \yH),
  \\[2\jot]
  \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}.
\end{gather}
The latter conditional probability is a hypergeometric distribution,
typical of \enquote{drawing without replacement} problems. The
combinatorial proof of the formulae above is in fact the same as for this
class of problems
\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

@@

Our initial knowledge (ignorance) $\yH$ has two important properties
derivable from the formulae above.

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-17T10:31:16+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

We have a population of $N$ binary neurons. We assume that they can be
distinguished, by their spike shapes for example; but other details, like
their locations, are unknown. The neurons have a joint state
$(X_1,\dotsc,X_N) \defs \yX$ having fixed but unknown binary values
$(R_1,\dotsc,R_N)\defs \yR \in \set{0,1}^N$. A particular sample of $n$
neurons from this population has joint state $(x_1, \dotsc, x_n) \defs \yx$
having fixed binary values $(r_1, \dotsc, r_n)\defs\yr$. We will consider
various averages of the population and the sample. For this purpose we
introduce a general averaging operator $\sav{\dotv}$ defined by
\begin{align}
  \av{\yX}  &\defd \tfrac{1}{N} (X_1 + X_2 + \dotsb + X_N),
%  &
%  \sav{\yx}  &\defd \tfrac{1}{n} (x_1 + x_2 + \dotsb + x_n),
    \\[2\jot]
  \av{\yX \yX} &\defd
\tbinom{N}{2}^{-1} (X_1 X_2 + X_1 X_3  + \dotsb +  X_{N-1} X_N),
%  &
%    \sav{\yx \yx} &\defd\tbinom{n}{2}^{-1} (x_1 x_2 + x_1 x_3  + \dotsb +  x_{n-1} x_n),
\\[2\jot]
\av{\yX\yX\yX} &\defd
                  \tbinom{N}{3}^{-1}
                  (X_1 X_2 X_3 +% X_1 X_2 X_4
                   \dotsb + X_{N-2} X_{N-1}X_N),
%  &
%    \sav{\yx\yx\yx} &\defd
%                      \tbinom{n}{3}^{-1} (x_1 x_2 x_3 + %x_1 x_2 x_4
%                       \dotsb + x_{n-2} x_{n-1} x_n),
\end{align}
and so on
These formulae say that $\yXf$ is the fraction of active neurons, $\yXXf$
the fraction of simultaneously active pairs out of all $\binom{N}{2}$
pairs, $\av{\yX\yX\yX}$ the fraction of simultaneously active triplets, and
so on. Products of states like $X_i \dotsm X_j$ also have values in
$\set{0,1}$; from this it can combinatorially be proved that
\begin{equation}
  \label{eq:products_intermsof_average}
  \av{\underbrace{\yX\dotsm\yX}_{\text{$m$ factors}}}
  = \binom{N}{m}^{-1}\binom{N\yXf}{m}.
\end{equation}
Analogous formulae  hold for quantities like $\yx$, $\yRf$, $\yrs$.

@@

By marginalization we obtain the probability for the state of the sample:
\begin{gather}
  \label{eq:marginal}
  \p(\yx = \yr \| \yH) = \binom{n}{n\yrs}^{-1} \p(\yxs=\yrs \| \yH),
  \\
  \shortintertext{with}
  \label{eq:subpop_average}
  \p(\yxs=\yrs \| \yH) = \sum_{N \yRf =0}^{N}
    \p(\yxs = \yrs \|\yXf=\yRf, \yH)\,
  \p(\yXf=\yRf \| \yH),
  \\[2\jot]
  \p(\yxs = \yrs \|\yXf=\yRf, \yH)=
  \binom{n}{n\yrs}\binom{N-n}{N \yRf-n\yrs}\binom{N}{N \yRf}^{-1}.
\end{gather}

The latter conditional probability is a hypergeometric distribution,
typical of \enquote{drawing without replacement} problems. The
combinatorial proof of the formulae above is in fact the same as for this
class of problems
\cites[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
This conditional probability acts a

@@

\begin{figure}[!t]
\centering
%\includegraphics[width=0.99\linewidth]{pop_sample_projection.pdf}\\
\includegraphics[width=0.99\linewidth]{pop_sample_projection2.pdf}%
\caption{hyperg}
\label{fig:hypergeom_proj}
\end{figure}%hypergeometric_identity.nb
% \end{wrapfigure}
Our initial knowledge (ignorance) $\yH$ has two important properties
derivable from the formulae above.
